Simulating Single Cell experiments from ScPCA data

### 2024-04-05 analyses/simulate-sce/notebooks/splatter-ScPCA.Rmd

suffixes <- c( "_processed.rds", "_filtered.rds", "_unfiltered.rds")

@@

# reduce the cell type data matrices, if present
if (!is.null(metadata(sce_sim)$singler_results)) {
  metadata(sce_sim)$singler_results <- metadata(sce_sim)$singler_results[cell_subset,]
}
if (!is.null(metadata(sce_sim)$cellassign_predictions)) {
  metadata(sce_sim)$cellassign_predictions <- metadata(sce_sim)$cellassign_predictions[cell_subset,]
}
```

@@

```{r}
random_label<- function(label_set, n){
  # randomly select n labels from the label set, ensuring each label is included at least once
  # ensure labels are unique
  label_set <- unique(label_set)
  if (length(label_set) > n) {
    stop("The number of labels must not be greater than the number requested")
  }
  r_labels <- sample(label_set, n, replace = TRUE)
  # add each label at least once
  idx <- sample.int(n, length(label_set))
  r_labels[idx] <- label_set
  return(r_labels)
}

```

@@

# create a mapping for singler ontology and annotation
singler_df <- colData(sce)|>
  data.frame()|>
  dplyr::select(singler_celltype_ontology, singler_celltype_annotation) |>
  dplyr::distinct()

@@

counts(sce_sim, withDimnames = FALSE) <- splatter::splatSimulate(sim_params)|>
  counts()
# recalculate dimension reduction
sce_sim <- sce_sim |>
  scater::logNormCounts()|>
  scater::runPCA(10) |> # we don't need all the PCA components
  scater::runUMAP(dimred = "PCA")
```

@@

# remove stats that will be recalculated
remove_stats <- c(
  "sum",
  "detected",
  "total",
  "subsets_mito_sum",
  "subsets_mito_detected",
  names(colData(sce_sim))[grep("altExp_.*_(sum|detected|percent)", names(colData(sce_sim)))]
)
colData(sce_sim)[,remove_stats] <- NULL

@@

# restore column order
colData(sce_sim) <- colData(sce_sim)[,colData_names]

```

>>>>

suffixes <- c("_processed.rds", "_filtered.rds", "_unfiltered.rds")

@@

# reduce the cell type data matrices, if present
if (!is.null(metadata(sce_sim)$singler_results)) {
  metadata(sce_sim)$singler_results <- metadata(sce_sim)$singler_results[cell_subset, ]
}
if (!is.null(metadata(sce_sim)$cellassign_predictions)) {
  metadata(sce_sim)$cellassign_predictions <- metadata(sce_sim)$cellassign_predictions[cell_subset, ]
}
```

@@

```{r}
random_label <- function(label_set, n) {
  # randomly select n labels from the label set, ensuring each label is included at least once
  # ensure labels are unique
  label_set <- unique(label_set)
  if (length(label_set) > n) {
    stop("The number of labels must not be greater than the number requested")
  }
  r_labels <- sample(label_set, n, replace = TRUE)
  # add each label at least once
  idx <- sample.int(n, length(label_set))
  r_labels[idx] <- label_set
  return(r_labels)
}
```

@@

# create a mapping for singler ontology and annotation
singler_df <- colData(sce) |>
  data.frame() |>
  dplyr::select(singler_celltype_ontology, singler_celltype_annotation) |>
  dplyr::distinct()

@@

counts(sce_sim, withDimnames = FALSE) <- splatter::splatSimulate(sim_params) |>
  counts()
# recalculate dimension reduction
sce_sim <- sce_sim |>
  scater::logNormCounts() |>
  scater::runPCA(10) |> # we don't need all the PCA components
  scater::runUMAP(dimred = "PCA")
```

@@

# remove stats that will be recalculated
remove_stats <- c(
  "sum",
  "detected",
  "total",
  "subsets_mito_sum",
  "subsets_mito_detected",
  names(colData(sce_sim))[grep("altExp_.*_(sum|detected|percent)", names(colData(sce_sim)))]
)
colData(sce_sim)[, remove_stats] <- NULL

@@

# restore column order
colData(sce_sim) <- colData(sce_sim)[, colData_names]
```
